---
title: "Manuscript title"
author:
    - name: "Dylan Russell, M.D."
      url: https://www.github.com/dylanrussellmd
      affiliation: Tripler Army Medical Center
      orcid_id: 0000-0002-9543-9897
date: |
  | "`r format(Sys.time(), '%B %d, %Y')`"
output:
  distill::distill_article:
    toc: true
    toc_depth: 1
    default: true
  bookdown::word_document2:
    reference_docx: "reference.docx"
    number_sections: false
bibliography: bibliography.bib
csl: surgical-endoscopy.csl
---

```{r preamble, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  include = FALSE
)
```

```{r libraries}
# Add library(this_package)
library(manuscriptr)
library(magrittr)
library(kableExtra)
library(ggplot2)
```

```{r results, child='results.Rmd'}
# Edit results code in "results.Rmd". 
# Written separately to keep code logic out of manuscript.
```

# Abstract {.unlisted .unnumbered}

**Introduction**:

**Methods**:

**Results**:

**Conclusion**:

# Introduction


# Methods

## Data Source and Study Population

## Outcomes

## Statistical Analysis


# Results

```{r figstabs, child='figstabs.Rmd'}
# Edit all figures and tables in "figstabs.Rmd".
# Written separately to keep code logic out of manuscript.
```

# Discussion


# Conclusion


# Disclosures {.unlisted .unnumbered}

The views and opinions expressed in this article are those solely of the authors
and in no way reflect the opinions of the United States Government. The authors 
have no conflicts of interest or relevant financial ties to disclose.

```{r citations, include = TRUE}
# If Word document output, needs a References section header
if(knitr::opts_knit$get("rmarkdown.pandoc.to") == "docx") {
  knitr::asis_output("# References")
}
```
